---
title: "Fuzzing stuff"
author: "Chris Wong"
date: "December 9, 2015"
output: html_document
---

```{r, echo=FALSE}
library(ggplot2)
```

```{r, echo=FALSE}
fuzz <- read.csv('fuzz.csv')
names(fuzz)[names(fuzz) == 'id'] <- 'size'

fuzz.lm <- coef(with(fuzz, lm(Document..updateStyle ~ size)))
fuzz.cor <- with(fuzz, cor(size, Document..updateStyle))
```

For the first experiment, the author generated random DOM fragments, and measured the time taken to call `Document::updateStyle`. `r length(fuzz$size)` samples were collected in total.

Update times fit a linear regression model with intercept `r format(fuzz.lm[1])` and slope `r format(fuzz.lm[2])`. The correlation coefficient is `r format(fuzz.cor)`, suggesting a very strong correlation.

```{r, echo=FALSE}
ggplot(fuzz, aes(x = size, y = Document..updateStyle)) +
  labs(title = "Passing random DOM fragments to Document::updateStyle",
       x = "Length of DOM string (bytes)",
       y = "Time taken in Document::updateStyle (μs)") +
  geom_point() +
  geom_smooth(method = 'lm')
```