---
title: "Style engine performance experiments"
author: "Chris Wong"
date: "December 16, 2015"
output:
  html_document:
    keep_md: yes
---

```{r, echo=FALSE}
library(ggplot2)
```

```{r, echo=FALSE}
fuzz <- read.csv('fuzz1/summary.csv')
names(fuzz)[names(fuzz) == 'id'] <- 'size'

fuzz.lm <- coef(with(fuzz, lm(Document..updateStyle ~ size)))
fuzz.cor <- with(fuzz, cor(size, Document..updateStyle))
```

For the first experiment, the author generated random DOM fragments, and measured the time taken to call `Document::updateStyle`. `r length(fuzz$size)` samples were collected in total.

Update times fit a linear regression model with intercept `r format(fuzz.lm[1])` and slope `r format(fuzz.lm[2])`. The correlation coefficient is `r format(fuzz.cor)`, suggesting a very strong correlation.

```{r, echo=FALSE}
ggplot(fuzz, aes(x = size, y = Document..updateStyle)) +
  labs(title = "Passing random DOM fragments to Document::updateStyle",
       x = "Length of DOM string (bytes)",
       y = "Time taken in Document::updateStyle (μs)") +
  geom_point() +
  geom_smooth(method = 'lm')
```

# Varying branching factor

```{r, echo=FALSE}
fuzz2.branchiness <- read.csv('fuzz2/summary-branchiness.csv')
names(fuzz2.branchiness)[names(fuzz2.branchiness) == 'id'] <- 'branchiness'
```

In the next experiment, the HTML fragments had a flat structure with no nesting. `r length(fuzz2.branchiness$branchiness)` samples were collected in total.

As before, the update time increased linearly with the number of nodes.

```{r, echo=FALSE}
ggplot(fuzz2.branchiness, aes(x = branchiness, y = Document..updateStyle)) +
  labs(title = "Random DOM fragments by branching factor",
       x = "Branching factor",
       y = "Time taken in Document::updateStyle (μs)") +
  geom_point() +
  geom_smooth(method = 'lm')
```

# Varying tree height

```{r, echo=FALSE}
fuzz2.depthicity <- read.csv('fuzz2/summary-depthicity.csv')
names(fuzz2.depthicity)[names(fuzz2.depthicity) == 'id'] <- 'depthicity'
```

In the final experiment, the fragments had a deeply nested structure. All nodes, except the innermost leaf node, had exactly one child. `r length(fuzz2.depthicity$depthicity)` samples were collected in total.

Again, the growth was linear.

```{r, echo=FALSE}
ggplot(fuzz2.depthicity, aes(x = depthicity, y = Document..updateStyle)) +
  labs(title = "Random DOM fragments by tree height",
       x = "Tree height",
       y = "Time taken in Document::updateStyle (μs)") +
  geom_point() +
  geom_smooth(method = 'lm')
```